---
title: "Homework1"
author: "Sarahy Martinez"
date: "2024-09-16"
output: github_document
---


# Problem 1

Load the dataset

```{r download_data}

data("penguins", package = "palmerpenguins")

```


Load library to allow for pull function, display the dataset, ,get the number of rows and columns, find the mean flipper length


```{r penguins_dataset,message= FALSE}

library(dplyr)

summary (penguins)

penguins_rows = nrow(penguins)
penguins_column = ncol(penguins)
 
mean(pull(penguins, flipper_length_mm), na.rm = TRUE)

     
```

In the dataset we can find the species, island, bill length, bill depth, flipper length, body mass, sex, and year. 
There are a total of 152 Adelie, 68 Chinstrap, and 124 Gentoo. There are also 165 females and 168 males. The 
size of the dataset is `r nrow(penguins)` and `r ncol(penguins)`. The mean flipper length is also `mean(pull(penguins, flipper_length_mm))`.


Creating the scatterplot for penguins, load libraries to use function for plot formation and ggsave to save plot 

```{r penguin_scatter, message= FALSE}

library(tidyverse)
library(ggplot2)
library(palmerpenguins)

ggplot(drop_na(penguins), aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguins_scatter.pdf")

```


# Problem 2

creating dataframe and calculating means 
```{r dataframe, message=FALSE}
problem_2 = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp >0,
  character_vec = c("A", "B", "C", "D", "E", "F", "G","I", "J", "K"),
  factor_vector =  factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))
)

means <- tibble(
  norm_samp_mean = mean(pull(problem_2, norm_samp)),
  norm_samp_pos_mean = mean(pull(problem_2, norm_samp_pos)),
  character_vec_mean = mean(pull(problem_2, character_vec)),
  factor_vector_mean = mean(pull(problem_2, factor_vector)),
)

```



