---
title: "Homework1"
author: "Sarahy Martinez"
date: "2024-09-16"
output: github_document
---


# Problem 1

Load the dataset

```{r download_data}

data("penguins", package = "palmerpenguins")

```


Load library to allow for pull function, display the dataset, ,get the number of rows and columns, find the mean flipper length


```{r penguins_dataset,message= FALSE}

library(dplyr)

summary (penguins)

penguins_rows = nrow(penguins)
penguins_column = ncol(penguins)
 
mean(pull(penguins, flipper_length_mm), na.rm = TRUE)

     
```

In the dataset we can find the species, island, bill length, bill depth, flipper length, body mass, sex, and year. 
There are a total of 152 Adelie, 68 Chinstrap, and 124 Gentoo. There are also 165 females and 168 males. The 
size of the dataset is `r nrow(penguins)` and `r ncol(penguins)`. The mean flipper length is also `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)`.


Creating the scatterplot for penguins, load libraries to use function for plot formation and ggsave to save plot 

```{r penguin_scatter, message= FALSE}

library(tidyverse)
library(ggplot2)
library(palmerpenguins)

ggplot(drop_na(penguins), aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()

ggsave("penguins_scatter.pdf")

```


# Problem 2

creating dataframe and calculating means 
```{r dataframe, eval=FALSE}
problem_2 = tibble(
  norm_samp = rnorm(10),
  norm_samp_pos = norm_samp >0,
  character_vec = c("A" ,"B" ,"C","D" ,"E" ,"F" ,"G" ,"I" ,"J" ,"K"),
  factor_vector =  factor(sample(c("Level1", "Level2", "Level3"), 10, replace = TRUE))
)

means <- tibble(
  norm_samp_mean = mean(pull(problem_2, norm_samp)),
  norm_samp_pos_mean = mean(as.numeric(pull(problem_2, norm_samp_pos))),
  character_vec_mean = mean(as.numeric(pull(problem_2, character_vec))),
  factor_vector_mean = mean(as.numeric(pull(problem_2, factor_vector))),
)

```

What works when writing a code chunk that applies the as.numeric function to the logical, character, and factor variable are the logical and factor variable but not the character vector. The "as.numeric" function is used to convert non-numerical values to numerical values. As a result, we are able to convert factors and logical vectors to numerical, however character cannot be transformed possibly due to formatting or we need to apply as.character to the non-numeric characters.
